The Evolution of Phenotypically Invariant Genetic Networks

### 2017-11-16 kalman_walk_paper.tex
added Haldane's rule. edited the discussion.

\paragraph{A new mechanism underlying Haldane's rule}
The faster rate of fitness decay observed in $F_2$ relative to $F_1$ hybrids is consistent with Haldane's rule: the observation that if only one hybrid sex is sterile or inviable it is likely the heterogametic sex (e.g. the male XY) \citep{haldane1922sex}.
If the regulatory coefficients for one system are shared between the sex and one or more autosomal chromosomes, and dosage compensation mechanisms are present (e.g. the homogametic sex silences or inactivates the expression of one of her two copies), $F_1$ males are effectively equivalent to purely autosomal-system $F_2$s. This is because incompatibilities between loci on the sex chromome with autosomal loci in $F_1$ males are less likely to be balanced.
\footnote{For example, in a 2 gene network where the first gene resides on an autosome and the second gene on the X chromosome, the male with system $\left( \left[ \begin{smallmatrix} a_1 & a_2 \end{smallmatrix} \right], \left[ \begin{smallmatrix} a_1 & a_2 \\ X_1 & X_2 \end{smallmatrix} \right] \right)$ hybridizing with the homozygous female $\left( \left[\begin{smallmatrix} \bar a_1 & \bar a_2 \\ \bar X_1 & \bar X_2 \end{smallmatrix}\right] \right)$ will produce an $F_1$ male offspring with dynamics given by $ \left[\begin{smallmatrix} 0.5(a_1 + \bar a_1) & 0.5(a_2 + \bar a_2) \\ X_1 & X_2 \end{smallmatrix}\right]$. If both genes resided on the autosomes this system would only be possible in an $F_2$ cross.}

>>>>

\paragraph{A new mechanism underlying Haldane's rule}
The faster rate of fitness decay observed in $F_2$ relative to $F_1$ hybrids is consistent with Haldane's rule: the observation that if only one hybrid sex is sterile or inviable it is likely the heterogametic sex (e.g. the male XY) \citep{haldane1922sex}.
If the regulatory coefficients for one system are shared between the sex and one or more autosomal chromosomes, and dosage compensation mechanisms are present (e.g. the homogametic sex silences or inactivates the expression of one of her two copies), $F_1$ males are effectively equivalent to purely autosomal-system $F_2$s. This is because incompatibilities between loci on the sex chromome with autosomal loci in $F_1$ males are less likely to be balanced.
\footnote{For example, in a 2 gene network where the first gene resides on an autosome and the second gene on the X chromosome, the male with system $\left( \left[ \begin{smallmatrix} a_1 & a_2 \end{smallmatrix} \right], \left[ \begin{smallmatrix} a_1 & a_2 \\ X_1 & X_2 \end{smallmatrix} \right] \right)$ hybridizing with the homozygous female $\left( \left[\begin{smallmatrix} \bar a_1 & \bar a_2 \\ \bar X_1 & \bar X_2 \end{smallmatrix}\right] \right)$ will produce an $F_1$ male offspring with dynamics given by $ \left[\begin{smallmatrix} 0.5(a_1 + \bar a_1) & 0.5(a_2 + \bar a_2) \\ \bar X_1 & \bar X_2 \end{smallmatrix}\right]$. If both genes resided on the autosomes this system would only be possible in an $F_2$ cross.}
